---
title: "R for data science"
author: "Maruf Ahmed Bhuiyan"
date: "5/1/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(tidyverse)
mpg

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point()

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

```

# 3.2.4 Exercises
+ Run ggplot(data = mpg). What do you see? 
```{r}
ggplot(data = mpg)
```

+ How many rows are in mpg? How many columns?
```{r}
nrow(mpg)
ncol(mpg)
```

+ What does the drv variable describe? Read the help for ?mpg to find out.
  Type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd
+ Make a scatterplot of hwy vs cyl.
```{r}
ggplot(data = mpg) +
        geom_point(mapping = aes(x = hwy, y = cyl))
```

+ What happens if you make a scatterplot of class vs drv? Why is the plot not useful?
```{r}
ggplot(data = mpg ) + 
          geom_point (mapping = aes(x = class , y = drv))
```

```{r}
ggplot(data = mpg ) + 
          geom_point (mapping = aes(x = displ, y = hwy, 
                                    color = class))
```


```{r}
ggplot(data = mpg ) + 
          geom_point (mapping = aes(x = displ, y = hwy, 
                                    color = class, size = class))
```

```{r}
ggplot(data =  mpg) + 
          geom_point (mapping = aes(x = displ , y =hwy,
                                    color = class, 
                                    size = class,
                                    shape = class))
```

```{r}
ggplot(data =  mpg) + 
          geom_point (mapping = aes(x = displ , y =hwy,
                                    color = class, 
                                    size = class,
                                    shape = class,
                                    alpha = class))
```

```{r}
ggplot(data =  mpg) + 
          geom_point (mapping = 
                        aes(x = displ , y =hwy,
                        size = class,
                        shape = class,
                        alpha = class), color = "blue")
```

# 3.3.1 Exercises
+ What’s gone wrong with this code? Why are the points not blue?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```
### Manual mapping of aesthetics must be specified outside the ___aes___ argument.
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

+ Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?
```{r}
?mpg
str(mpg)
mpg
```
### Since, __mpg__ is a tibble, simply printing it would tell me about the variables.
### We can also use __str()__ to view how the variables are coded.
### No categorical variable is present in this tibble. displ,  year, cyl, cty, hwy are continuous.
+ Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?
```{r}
#ggplot(data =  mpg) + 
          #geom_point (mapping = aes(x = displ, y = hwy, 
                                    #size = cyl, 
                                    #color = cyl,
                                    #shape = cyl))
```
### Continuous variables can not be mapped to shape.
+ What happens if you map the same variable to multiple aesthetics?
```{r}
ggplot(data =  mpg) + 
          geom_point (mapping = aes(x = displ , y =hwy,
                                    color = class, 
                                    size = class,
                                    shape = class,
                                    alpha = class))
```

+ What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)
```{r}
?geom_point

# For shapes that have a border (21-24), we can color the inside (fill()) and
# outside(border-color()) separately. The stroke aesthetic can be used to modify the width of the border

# The hollow shapes (0–14) have a border determined by color; 
# The solid shapes (15–18) are filled with colour; 
# The filled shapes (21–24) have a border of colour and are filled with fill.

ggplot(mpg, aes(displ, hwy)) +
  geom_point(shape = 21, colour = "blue", fill = "cyan3", size = 1, stroke = 4)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(shape = 22, colour = "blue", fill = "cyan3", size = 2, stroke = 3)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(shape = 23, colour = "blue", fill = "cyan3", size = 3, stroke = 2)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(shape = 24, colour = "blue", fill = "cyan3", size = 4, stroke = 1)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(shape = 24, colour = "blue", fill = "cyan3", size = 5, stroke = 0.5)

```
+ What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y.
```{r}
#ggplot(mpg) +
  #geom_point(mapping = aes(displ, hwy), shape = 24, 
             #colour = displ < 5, fill = "cyan3", 
             #size = 10, stroke = 0.5)
```

```{r}
#ggplot(data = mpg) 
#       + geom_point(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ cyl, nrow = 2)

table(mpg$manufacturer)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ manufacturer, nrow = 5)

table(mpg$model)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ model, nrow = 10)

table(mpg$trans)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ trans, nrow = 5)
```


```{r}
ggplot(data = mpg ) + 
          geom_point(mapping = aes(x = displ, y =hwy)) +
          facet_grid(cyl~drv)

ggplot(data =  mpg) + 
          geom_point(mapping = aes(x = displ, y = hwy)) +
          facet_grid(.~drv)
ggplot(data =  mpg) + 
          geom_point(mapping = aes(x = displ, y = hwy)) +
          facet_grid(~drv)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(.~ class)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(~ class)
```

# 3.5.1 Exercises
+ What happens if you facet on a continuous variable?
```{r}
ggplot(data = mpg ) + 
          geom_point(mapping = aes(x = hwy, y = manufacturer)) +
          facet_wrap(~ displ)

ggplot(data = mpg ) + 
          geom_point(mapping = aes(x = displ, y = hwy)) +
          facet_wrap(~ cty)
table(mpg$cty)
```
+ What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl)) +
  facet_grid(drv~cyl)

### Empty plots mean there is no car to match this configuration. For example, there is no 4WD car that has a 5 cylinder engine which nicely correlates in both graphs.
```

+ What plots does the following code make? What does . do?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)

###. (dot) represents how do i want to graph to orient. For example, ___facet_grid(drv ~ .)___ means i want facet_grid to arrange the graphs in rows. Accordingly, ___facet_grid(. ~ cyl)___ instructs to arrange in columns.
```

+ Take the first faceted plot in this section: What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)

###Faceting makes it easy to understand because of descrete graphs compared to colors.  
###For larger datasets, faceting may make the graph too cramped.
```

+ Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol arguments?
```{r}
?facet_wrap
?facet_grid

# facet_wrap wraps a 1d sequence of panels into 2d. 
# This is generally a better use of screen space than facet_grid() 
# because most displays are roughly rectangular.

# facet_grid() forms a matrix of panels defined by row and column faceting variables. 
# It is most useful when you have two discrete variables, and all combinations of the 
# variables exist in the data.

### nrow and ncol represents number of rows and columns respectively.
### scales, shrink, labeller, switch are other options to control the layout of individual panels. 
### ___facet_grid()___ forms a matrix, that's why it doesn't have nrow and ncol arguments ??
```

+ When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(class ~ drv)

### Putting variables with many levels in the rows shrinks the space like the above graph.
### That's why it recommended to put the variable with more unique levels in the columns.
```

```{r}
ggplot(data =  mpg) + 
          geom_smooth(mapping = aes(x = displ , y = hwy))

ggplot(data =  mpg) + 
          geom_point(mapping = aes(x = displ , y = hwy))

ggplot(data =  mpg) + 
          geom_smooth(mapping = aes(x = displ , y = hwy, linetype = drv))

# ggplot(data =  mpg) + 
         # geom_smooth(mapping = aes(x = displ , y = hwy, linetype = class))
```

```{r}
ggplot(data =  mpg) + 
          geom_smooth(mapping = aes(x = displ , y = hwy, linetype = drv))

ggplot(data =  mpg) + 
          geom_smooth(mapping = aes(x = displ , y = hwy, group = drv))

ggplot(data =  mpg) + 
          geom_smooth(mapping = aes(x = displ , y = hwy, color = drv))

ggplot(data =  mpg) + 
          geom_smooth(mapping = aes(x = displ , y = hwy, color = drv), show.legend = FALSE)
```

```{r}
ggplot(data =  mpg) + 
          geom_smooth(mapping = aes(x = displ , y = hwy)) +
          geom_point(mapping = aes(x = displ, y = hwy))

ggplot(data =  mpg) + 
          geom_smooth(mapping = aes(x = displ , y = hwy, linetype = drv)) +
          geom_point(mapping = aes(x = displ, y = hwy))
```

```{r}

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth()

```

```{r}
table(mpg$class)
data = filter(mpg, class == "suv")

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth(data = data, se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class)) + 
  geom_point() + 
  geom_smooth(data = data, se = FALSE)
```

# 3.6.1 Exercises
1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?
```{r}
# Line Chart <- geom_line
# Boxplot <- geom_boxplot
# Histogram <- geom_histogram
# Area chart <- geom_density
```

2. Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```
3. What does show.legend = FALSE do? What happens if you remove it? Why do you think I used it earlier in the chapter?
```{r}
# Hides legend
# It's true by defualt. So, legend will be shown.
# To hide legend
```
4. What does the se argument to geom_smooth() do?
```{r}
# Creates a smooth (loess)line.
```
5. Will these two graphs look different? Why/why not?
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))

# No, they ara the same. In the first one the arguments will be used as global.
# In the second graph the same arguments are employed at local level.
```
6. Recreate the R code necessary to generate the following graphs.
```{r}
#install.packages("gridExtra")
#install.packages("cowplot")
#library(cowplot)

g1 <- ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
          geom_point(color = "black") + 
          geom_smooth(color = "blue", se = FALSE, show.legend = FALSE)

g2 <- ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
          geom_point(color = "black") + 
          geom_smooth(mapping = aes(group = drv), color = "blue", se = FALSE)

g3 <- ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
          geom_point() + 
          geom_smooth( se = FALSE)

g4 <- ggplot(data = mpg, mapping = aes(x = displ , y =hwy)) + 
          geom_point(mapping = aes(color = drv)) +
          geom_smooth(se = FALSE)

g5 <- ggplot(data = mpg, mapping = aes( x = displ, y = hwy, color = drv) ) + 
          geom_point() +
          geom_smooth(mapping = aes(linetype = drv), se = FALSE)

g6 <- ggplot(data = mpg, mapping = aes( x = displ, y = hwy)) + 
          geom_point(mapping = aes(color = drv)) +
          geom_point(shape = 21, color = 'white', size = 2, stroke = 1)
                     
theme_set(theme_bw())
plot_grid(g1,g2,g3,g4,g5,g6, labels = c("1", "2", "3", "4", "5", "6"))

```

```{r}
ggplot(data =  mpg ) + 
          geom_bar(mapping = aes(x = class))

diamonds
dim(diamonds)
names(diamonds)
head(diamonds)

b1 <- ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))

s1 <- ggplot(data = diamonds) + 
          stat_count(mapping = aes(x = cut))

b2 <-ggplot(data = diamonds) + 
          geom_bar(mapping = aes(x = clarity))

s2 <- ggplot(data = diamonds) + 
          stat_count(mapping = aes(x = clarity))

plot_grid(b1, s1, b2, s2, labels = c("Bar1", "Count1", "Bar2", "Count2"), 
          nrow = 2, ncol = 2)

```

# 3.7.1 Exercises
1. What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?
```{r}
# Boxplot

ggplot(data = diamonds) + 
  stat_summary(mapping = aes(x = cut, y = depth),
               fun.min = min,
               fun.max = max,
               fun = median)

ggplot(data = diamonds) +
  geom_pointrange(mapping = aes(x = cut, y = depth),
                  fun.min = min,
                  fun.max = max,
                  fun = median,
                  stat = "summary")

```

2. What does geom_col() do? How is it different to geom_bar()?
```{r}
# There are two types of bar charts: geom_bar() and geom_col(). geom_bar() makes 
# the height of the bar proportional to the number of cases in each group (or if 
# the weight aesthetic is supplied, the sum of the weights). If you want the 
# heights of the bars to represent values in the data, use geom_col() instead.
# geom_bar() uses stat_count() by default: it counts the number of cases at each x # position. geom_col() uses stat_identity(): it leaves the data as is.


```

3. Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?
|       geom          |         stat        |
|---------------------|---------------------|
| `geom_bar()`        | `stat_count()`      |
| `geom_bin2d()`      | `stat_bin_2d()`     |
| `geom_boxplot()`    | `stat_boxplot()`    |
| `geom_contour()`    | `stat_contour()`    |
| `geom_count()`      | `stat_sum()`        |
| `geom_density()`    | `stat_density()`    |
| `geom_density_2d()` | `stat_density_2d()` |
| `geom_hex()`        | `stat_hex()`        |
| `geom_freqpoly()`   | `stat_bin()`        |
| `geom_histogram()`  | `stat_bin()`        |
| `geom_qq_line()`    | `stat_qq_line()`    |
| `geom_qq()`         | `stat_qq()`         |
| `geom_quantile()`   | `stat_quantile()`   |
| `geom_smooth()`     | `stat_smooth()`     |
| `geom_violin()`     | `stat_violin()`     |
| `geom_sf()`         | `stat_sf()`         |

4. What variables does stat_smooth() compute? What parameters control its behaviour?
The function  stat_smooth() computes linear regresssion model (lm)
The parameters that control the behavior of `stat_smooth()` include
-   `method`: the method used to
-   `formula`: the formula are parameters such as `method` which determines which method is used to calculate the predictions and confidence interval, and some other arguments that are passed to that.
-   `na.rm`:
5. In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?
```{r}

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = stat(prop)))

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = stat(prop), group = 1))

# without group = 1, all proporitonsare calulated in each group which
# always equals to one.
# with group = 1, all proportions are calculated as a whole.

```


```{r}
ggplot(data =  diamonds) + 
          geom_bar(mapping = aes(x = cut, color = cut))

ggplot(data = diamonds ) + 
          geom_bar(mapping = aes(x = cut , fill = cut))

ggplot(data = diamonds ) + 
          geom_bar(mapping = aes(x = cut, fill = clarity))

ggplot(data = diamonds ) + 
          geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")

ggplot(data =  diamonds, mapping = aes(x = cut, fill = clarity)) + 
          geom_bar(position = "dodge")

ggplot(data =  diamonds, mapping = aes(x = cut, fill = clarity)) + 
          geom_bar(position = "identity")

ggplot(data =  diamonds, mapping = aes(x = cut, fill = clarity)) + 
          geom_bar( fill = NA, position = "identity")

ggplot(data =  diamonds, mapping = aes(x = cut, fill = clarity)) + 
          geom_bar( alpha = 0.2, position = "identity")

ggplot(data =  diamonds, mapping = aes(x = cut, fill = clarity)) + 
          geom_bar(position = "fill")

ggplot(data = mpg ) + 
          geom_point(mapping = aes(x = displ, y = hwy), position = "jitter")

ggplot(data = mpg ) + 
          geom_point(mapping = aes(x = displ, y = hwy))

ggplot(data = mpg ) +
          geom_jitter(mapping = aes(x = displ, y = hwy))

```

# 3.8.1 Exercises
1. What is the problem with this plot? How could you improve it?
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()

# The problem with this plot is that since there are many observations with 
# the same values of cty and hwy, many data points overlap with each other and it 
# fails to show where the mass of the data is. We can use geom_jitter() instead:

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_jitter()

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_jitter() +
  geom_smooth()
```
2. What parameters to geom_jitter() control the amount of jittering?
```{r}
?geom_jitter

# width	& height	
# Amount of vertical and horizontal jitter. The jitter is added in both positive and # negative directions, so the total spread is twice the value specified here.

```
3. Compare and contrast geom_jitter() with geom_count().
```{r}
# Both geom_jitter() and geom_count() can better represent the data when there 
# are many overlapping points and show where the mass of the data is. geom_jitter()
# achieves this by slightly moving the overlapping points vertically and horizontally,
# sacrificing statistical accuracy while more clearly in showing the overlapping
# points. geom_count() counts the overlapping points, and maps the counts to the size
# of the points instead.

ggplot(data = mpg) + 
          geom_jitter(mapping = aes(x = displ, y = hwy))

ggplot(data = mpg) + 
          geom_count(mapping = aes(x = displ, y = hwy))
```

4. What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.
```{r}
?geom_boxplot

# The default position for geom_boxplot is "dodge2"
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +
      geom_boxplot(mapping = aes(color = factor(year)))

```

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +
      geom_boxplot(mapping = aes(color = factor(year)))

ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +
      geom_boxplot(mapping = aes(color = factor(year))) +
      coord_flip()
```

```{r}
#install.packages("maps")
library(maps)

nz <- map_data("nz")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black") +
  coord_quickmap()
```

```{r}
bar <- ggplot(data = mpg) + 
       geom_bar(mapping = aes(x = hwy, fill = drv), 
                show.legend = FALSE, width = 1) + 
                theme(aspect.ratio = 1) +
                labs(x = NULL, y = NULL)
bar

bar + coord_flip()
bar + coord_polar()

bar <- ggplot(data = diamonds) + 
       geom_bar(mapping = aes(x = cut, fill = cut), 
                show.legend = FALSE,width = 1) + 
                theme(aspect.ratio = 1) +
                labs(x = NULL, y = NULL)
bar

bar + coord_flip()
bar + coord_polar()
```

# 3.9.1 Exercises
1. Turn a stacked bar chart into a pie chart using coord_polar().
```{r}
ggplot(data = mpg) + 
          geom_bar(mapping = aes(x = class, fill = class))

ggplot(data = mpg) + 
          geom_bar(mapping = aes(x = class, fill = class)) +
          coord_polar()
```

2. What does labs() do? Read the documentation.
```{r}
# labs() allows us to modify axis, plot, and legend labels.

ggplot(data = mpg) + 
          geom_bar(mapping = aes(x = class, fill = class)) +
          labs(x = "Class", y = "Total number", 
               title = "Title",
               subtitle = "Subtitle",
               caption = "Caption",
               tag = "Tag")
```

3. What’s the difference between coord_quickmap() and coord_map()?
```{r}
nz <- map_data("nz")

ggplot(nz, aes(long, lat, group = group)) +
                geom_polygon(fill = "white", colour = "black")

ggplot(nz, aes(long, lat, group = group)) +
                geom_polygon(fill = "white", colour = "black") +
                coord_quickmap()

ggplot(nz, aes(long, lat, group = group)) +
                geom_polygon(fill = "white", colour = "black") +
                coord_map()

if (require("maps")) {
nz <- map_data("nz")
# Prepare a map of NZ
nzmap <- ggplot(nz, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "white", colour = "black")

# Plot it in cartesian coordinates
nzmap
}

#install.packages("mapproj")
library(mapproj)

if (require("maps")) {
# With correct mercator projection
nzmap + coord_map()
}

if (require("maps")) {
# With the aspect ratio approximation
nzmap + coord_quickmap()
}

```

4. What does the plot below tell you about the relationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do?
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline()

# coord_fixed forces a specified aspect ratio between the physical representation 
# of the units on the axes. The ratio is 1 by default. It is important to fix 
# the aspect ratio in this case because hwy and cty are measured in the same unit
# (miles per gallon). Any other aspect ratios will give a visually incorrect
# representation and might lead us to believe that one increasese at a faster rate 
# than the other.
```

```{r}
seq(1,10,2)
(x <- "hello world")
(y <- seq(0,50, 5))

library(tidyverse)

#ggplot(do#ta = mpg) + 
  #geom_point(mapping = aes(x = displ, y = hwy))

#fl#iter(mpg, cyl =# 8)
#filter(diamond#, carat > 3)


ggplot(data = mpg) + 
        geom_point(mapping = aes(x = displ, y = hwy))

filter(mpg, cyl == 8)
filter(diamonds, carat > 3)

```

```{r}

#install.packages("nycflights13")

library(nycflights13)
library(tidyverse)

flights
dim(flights)

names(flights)

jan1 <- flights %>%
      filter(month == 1, 
             day == 1)

dim(jan1)

two <- sqrt(2) ^ 2
two # 2
two ==2 # FALSE
near(two, 2) # TRUE

nov_dec1 <- filter(flights, month == 11 | month == 12)

nov_dec2 <- filter(flights, month %in% c(11,12))

identical(nov_dec1, nov_dec2)

delay1 <- flights %>%
              filter(!(arr_delay > 120 | dep_delay > 120))
delay1
delay2 <- flights %>%
              filter( arr_delay <= 120 , dep_delay <= 120)
delay2

identical(delay2, delay2)
sum(is.na(flights))

```

# FILTER

1. Find all flights that
1. Had an arrival delay of two or more hours
```{r}
flights %>%
      filter( arr_delay >= 120)
```
2. Flew to Houston (IAH or HOU)
```{r}
d1 <- flights %>%
      filter(dest == "IAH" | dest == "HOU")

d2 <- flights %>%
      filter(dest %in% c("IAH", "HOU"))

identical(d1, d2)

```

3. Were operated by United, American, or Delta
```{r}
?flights

# United, American, Delta > two letter abbreviation UA, AA, DL

flights %>%
        filter(carrier %in% c('UA', 'AA', 'DL') )
```

4. Departed in summer (July, August, and September)
```{r}

flights %>%
        filter(month %in% c(7,8,9))
```

5. Arrived more than two hours late, but didn’t leave late
```{r}
flights %>%
      filter(arr_delay > 120 & dep_delay <= 0)
```

6. Were delayed by at least an hour, but made up over 30 minutes in flight
```{r}

flights %>%
        filter(dep_delay >= 60, dep_delay > arr_delay + 30)

```

7. Departed between midnight and 6am (inclusive)
```{r}

m1 <- flights %>%
      filter(dep_time >= 0, dep_time <= 600)

```

2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?
```{r}

m2 <- flights %>%
        filter(between(dep_time, 0 , 600))

identical(m1,m2)
```

3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
sum(is.na(flights$dep_time))

flights %>%
        filter(is.na(dep_time))
# These flights were most likely cancelled.
```

4. Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)
```{r}
NA ^0

NA | TRUE

FALSE & NA

NA * 0


# One way to think of NA is that it can be a placeholder for any possible values. 
# By this logic:
## any values raised to the power of 0 is 0
## NA | TRUE, anything else OR TRUE is always TRUE
## NA | FALSE, anything else AND FALSE is always FALSE

# In case of  NA * 0 should give 0 when NA is replaced by any number but Inf.
# Inf * 0 and -Inf * 0 both give NA.

```

# ARRANGE

```{r}
flights %>%
      arrange(year, month, day)

flights %>%
      arrange(desc(dep_delay))
```

# 5.3.1 Exercises
1. How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).
```{r}
flights %>%
      arrange(desc(is.na(dep_delay)))
```

2. Sort flights to find the most delayed flights. Find the flights that left earliest.
```{r}
flights %>%
        arrange(desc(dep_delay))

flights %>%
        arrange(dep_delay)
```

3. Sort flights to find the fastest (highest speed) flights.
```{r}
# speed = distance / time

flights %>%
        arrange(air_time, desc(distance))
```

4. Which flights travelled the farthest? Which travelled the shortest?
```{r}
distance <- flights %>%
        select(year, month, day, distance) %>%
        arrange(desc(distance))

head(distance)  # farthest
tail(distance)  # shortest

```


```{r}
flights %>%
        select(year, month, day, dep_delay) %>% 
        filter(month ==1 & day == 1)

flights %>%
        select( -(year:day))

flights %>%
        select((year:day))


flights %>%
        select(starts_with("y"))

flights %>%
        select(ends_with("delay"))

flights %>%
        select(contains("y"))

flights %>%
        select(Year = year, Month = month)

flights %>%
        rename(Year = year, Month = month, Day = day)

flights %>%
        select(origin, dest, air_time, distance, everything())

```

# 5.4.1 Exercises
1. Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.
```{r}
flights %>% 
        select(dep_time, dep_delay, arr_time, arr_delay)

flights %>% 
        select(contains("time"), contains("delay"))

flights %>% 
        select(ends_with("time"), ends_with("delay"))

flights %>% 
        select(starts_with("dep"), starts_with("arr"))

flights %>% 
        select(contains("_"))

```

2. What happens if you include the name of a variable multiple times in a select() call?
```{r}
flights %>% 
        select(year, month, year, month)
```

3. What does the one_of() function do? Why might it be helpful in conjunction with this vector?
```{r}

vars <- c("year", "month", "day", "dep_delay", "arr_delay")

flights %>% 
        select(one_of(vars))

```

4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?
```{r}

select(flights, contains("TIME"))

# conatains is not case sensitive

select(flights, contains("TIME", ignore.case = FALSE))

# we can change the default behaviour by ignore.case = FALSE

```

```{r}
flights %>% 
        mutate(gain = dep_delay - arr_delay, speed = distance/air_time * 60) %>% 
        select(month, day, gain, speed, air_time, distance)

flights %>% 
        transmute(gain = dep_delay - arr_delay, speed = distance/air_time * 60)

flights %>% 
        transmute(gain = dep_delay - arr_delay, speed = distance/air_time * 60) %>% 

```

# 5.5.2 Exercises
1. Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.
```{r}
flights %>% 
      mutate(dep_time_mins = dep_time %/% 100 * 60 + dep_time %% 100,
                  sched_dep_time_mins = sched_dep_time %/% 100 * 60 +
                    sched_dep_time %% 100) %>% 
      select(starts_with('dep_time'), starts_with('sched'))

```
2. Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?
```{r}
flights %>% 
        mutate(dif = arr_time - dep_time) %>% 
        select(air_time, dif)

# I was expecting air_time and dif to be the same. But, that's not the case
# because arr_time is not a continous number.

f1 <- flights %>% 
      mutate(arr_time_mins = arr_time %/% 100 * 60 + arr_time %% 100,
             dep_time_mins = dep_time %/% 100 * 60 + dep_time %% 100) %>% 
      mutate(flight_time = arr_time_mins - dep_time_mins) %>% 
      select(air_time, flight_time) 

sum(f1$air_time == f1$flight_time, na.rm = TRUE)

```

3. Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?
```{r}
flights %>% 
      select(dep_time, sched_dep_time, dep_delay)

# The relation is dep_time = sched_dep_time + dep_delay

```

4. Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().

```{r}
flights %>% 
      select(month, day, dep_delay) %>% 
      arrange(min_rank(desc(dep_delay)))

flights %>% 
      select(month, day, dep_delay) %>% 
      arrange(desc(dep_delay))

# Why use min_rank() then ?? Not sure
```

5. What does 1:3 + 1:10 return? Why?
```{r}
c1 <- 1:3 + 1:10
c1
# this is basically what happens
# two vectors are created vector1(1,2,3) and vector2 (1,2,3,4,5,6,7,8,9,10)
# vector1(1,2,3) is repeated until it is same length as vector 2
# 1,2,3,1,2,3,1,2,3,1

c2 <- c(1,2,3,1,2,3,1,2,3,1) + 1:10

c1
c2

```
6. What trigonometric functions does R provide?
```{r}
# cos(x)
# sin(x)
# tan(x)

# acos(x)
# asin(x)
# atan(x)
# atan2(y, x)

# cospi(x)
# sinpi(x)
# tanpi(x)
```

# SUMMARISE

```{r}
flights %>% 
      group_by(month, day) %>% 
      summarise(delay = mean(dep_delay, na.rm = TRUE))

by_dest <- group_by(flights, dest)
delay <- summarise(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
)
delay <- filter(delay, count > 20, dest != "HNL")

ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)

flights %>% 
      group_by(dest) %>% 
      summarise(count = n(), 
                dist = mean(distance, na.rm = TRUE),
                delay = mean(arr_delay, na.rm = TRUE)) %>%
      filter(count > 20, dest != "HNL") %>% 
      ggplot(mapping = aes(x = dist, y = delay)) +
      geom_point(aes(size = count), color = "cyan3", alpha = 1/3) +
      geom_smooth(se = FALSE)
  

```






